---
title: ""
author: ""
output: word_document
---


```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.width=10, fig.height=6, fig.align = "left") 
```


```{r echo=FALSE, message=FALSE, warning=FALSE}


# 1 packages ----
library(tidyverse)
library(RColorBrewer)
library(RODBC)
library(flextable)
library(lme4)
library(birdnames)
options(scipen = 999)
source("C:/Users/scott.jennings/Documents/Projects/R_general/utility_functions/numbers2words.R")
source("C:/Users/scott.jennings/Documents/Projects/HEP/HEP_data_work/HEP_code/HEP_utility_functions.R")
# source("C:/Users/scott.jennings/Documents/Projects/HEP/HEP_data_work/HEP_code/HEP_data_summary_functions.R")


par_code = 68
zyear = 2021
  
```



```{r echo=FALSE}


hepdata_location = "C:/Users/scott.jennings/Documents/Projects/HEP/HEP_data_work/HEP_data/HEPDATA.accdb"
# all these functions are in HEP_utility_functions.R
hep_start <- hep_from_access(hepdata_location)
hep_sites <- hep_sites_from_access(hepdata_location) %>% 
  filter(parent.code == par_code)


hep_2020 <- read.csv("C:/Users/scott.jennings/Documents/Projects/HEP/rawhep_to_HEPDATA/data/as_HEPDATA/HEPDATA_2020.csv") %>% 
  clean_hep() %>% 
  right_join(., dplyr::select(hep_sites, code, parent.code, site.name, parent.site.name, county, subregion))

hep_2021 <- read.csv("C:/Users/scott.jennings/Documents/Projects/HEP/rawhep_to_HEPDATA/data/as_HEPDATA/HEPDATA_2021.csv") %>% 
  clean_hep() %>% 
  right_join(., dplyr::select(hep_sites, code, parent.code, site.name, parent.site.name, county, subregion))


hep <- hep_start %>% 
  clean_hep() %>% 
  filter(peakactvnsts >= 0) %>% 
  right_join(., dplyr::select(hep_sites, code, parent.code, site.name, parent.site.name, county, subregion))  %>% # this is a filtering join to par_code
  rbind(hep_2020, hep_2021) %>% 
  trim_hep_columns(stage = FALSE) %>% 
  cut_never_nested() %>% 
  rename(alpha.code = species) %>% 
  mutate(site.name = as.character(site.name),
         common.name = translate_bird_names(alpha.code, "alpha.code", "common.name")) %>%  
  ungroup() %>% 
  select(code, year, alpha.code, common.name, everything())


subcolonies <- hep %>% 
  distinct(parent.code, parent.site.name, code, site.name)


colony_first_year <- hep %>% 
  group_by(site.name) %>% 
  summarize(first.year = min(year),
            last.year = max(year))

par_colony <- hep %>% 
  distinct(parent.site.name)

species_years <- hep %>% 
  filter(peakactvnsts > 0) %>% 
  group_by(site.name, common.name) %>% 
  summarise(first.year = min(year),
            last.year = max(year),
            num.years = n()) %>% 
  ungroup() %>% 
  mutate(common.name = factor(common.name, levels = levels(core4spp))) %>% 
  arrange(common.name, site.name)


rounded_start_year <- floor(min(species_years$first.year)/5) * 5
rounded_end_year <- ceiling(min(zyear)/5) * 5

spp_richness <- hep %>% 
  distinct(common.name) %>% 
  summarise(spp.richness = n()) %>% 
  mutate(spp.richness = numbers2words(spp.richness))

```
Heron and egret nest monitoring for `r par_colony$parent.site.name`, `r zyear` season.
\
\
Audubon Canyon Ranch has monitored heron and egret nesting activity throughout the North San Francisco Bay Area since 1990. Each year, a ACR staff and citizen science volunteers monitor 60-80 active colony sites. Each colony is visited approximately 1-2 times per month between March-July. Observers collect information on which species are present and actively nesting, the number of nests of each species, how many nests survive to fledge, and how many young are produced per nests.  
\
\
ACR has monitored `r par_colony$parent.site.name` since `r min(colony_first_year$year)`. In that time, we have documented `r spp_richness$spp.richness` species nesting there (Table 1).
\
\
Table 1. Species detected nesting at `r par_colony$parent.site.name`, first year and number of years recorded nesting.
```{r echo=FALSE}
species_years %>% 
  select(-last.year) %>% 
  arrange(common.name) %>% 
  flextable() %>% 
  set_header_labels(site.name = "Subcolony",
                    common.name = "Species",
                    first.year = "First year\nrecorded nesting",
                    num.years = "Total years\nrecorded nesting") %>% 
  autofit() %>% 
  align_nottext_col(align = "center", header = TRUE, footer = TRUE) %>% 
  width(j = c(3, 4), width = 1)

```
\
\






Figure 1. Total number of nest at the `r par_colony$parent.site.name` colony, by species (and subcolony if applicable).
```{r echo=FALSE}

# peak active nests

num_nests <- hep %>% 
  select(common.name, site.name, year, peakactvnsts) %>% 
  group_by(site.name, common.name, year) %>% 
  summarise(total.nests = sum(peakactvnsts)) %>% 
  ungroup() %>% 
  mutate(common.name = factor(common.name, levels = levels(core4spp))) %>% 
  filter(total.nests > 0)


num_nests %>% 
  ggplot(group = site.name) +
  geom_col(aes(x = year, y = total.nests, fill = site.name)) +
  #geom_point(aes(x = year, y = total.nests)) +
  #geom_point(data = filter(num_nests, year == zyear), aes(x = year, y = total.nests), color = "red", size = 2) +
  #stat_smooth(aes(x = year, y = total.nests), color = "black", method = "lm", formula = y ~ x + I(x^2), se = FALSE) +
  scale_x_continuous(breaks = seq(rounded_start_year, rounded_end_year, by = 5), labels = seq(rounded_start_year, rounded_end_year, by = 5)) +
  facet_wrap(~common.name, scales = "free_y") +
  theme_bw() +
  ylab("Total nests") +
  xlab("Year") +
  theme(legend.title = element_blank()) 

```
\
\


Figure 2. Average number of chicks produced per successful nest at the `r par_colony$parent.site.name` colony, by species (and subcolony if applicable). This value is calculated for nests where brood size close to fledging was confidently counted. Error bar indicated standard error. Lack of error bar indicates only one nest for that species X year combination.
```{r echo=FALSE}

# number of chicks per nest


stage4_chicks <- hep %>% 
  select(year, common.name, site.name, contains("brd")) %>% 
  pivot_longer(contains("brd"), names_to = "brood.size", values_to = "num.nests") %>% 
  filter(num.nests > 0) %>% 
  mutate(brood.size = gsub("brd", "", brood.size),
         brood.size = as.numeric(brood.size)) %>% 
  filter(!is.na(num.nests))


chicks_per_nest <- stage4_chicks  %>% 
  group_by(year, site.name, common.name) %>% 
  summarise(total.stage4.chx = sum(brood.size * num.nests),
            stage4.nests = sum(num.nests)) %>% 
  ungroup() %>% 
  mutate(chicks.per.nest = total.stage4.chx/stage4.nests) %>% 
  ungroup() %>% 
  mutate(common.name = factor(common.name, levels = levels(core4spp))) %>% 
  arrange(common.name, year, site.name)

chx_per_nest_standard_error <- stage4_chicks %>% 
  full_join(chicks_per_nest) %>% 
  mutate(deviations = (brood.size - chicks.per.nest) * num.nests) %>% 
  group_by(year, site.name, common.name) %>% 
  summarise(sum.deviations = sum(deviations ^ 2),
            stage4.nests = sum(num.nests)) %>% 
  ungroup() %>% 
  mutate(st.deviation = sqrt(sum.deviations/(stage4.nests - 1)),
         st.error = st.deviation/sqrt(stage4.nests))


chicks_per_nest_se <- chx_per_nest_standard_error %>% 
  select(year, common.name, site.name, st.error) %>% 
  full_join(chicks_per_nest)

chicks_per_nest_se %>% 
  ggplot() +
  geom_col(aes(x = year, y = chicks.per.nest, fill = site.name)) +
  geom_errorbar(aes(x = year, y = chicks.per.nest, ymin = chicks.per.nest - st.error, ymax = chicks.per.nest + st.error)) +
  #geom_text(aes(x = year, y = -0.25, label = stage4.nests), size = 3) +
  #geom_point(aes(x = year, y = chicks.per.nest)) +
  #geom_point(data = filter(chicks_per_nest, year == zyear), aes(x = year, y = chicks.per.nest), color = "red", size = 2) +
  #stat_smooth(aes(x = year, y = chicks.per.nest), method = "lm", formula = y ~ 1, color = "black", se = FALSE) +
  facet_wrap(~common.name, scales = "free_y") +
  theme_bw() +
  ylab("Chicks/nest") +
  xlab("Year") +
  theme(legend.title = element_blank())

  


```
\
\

