---
title: "Heron and Egret nest monitoring summary"
output: word_document
---

```{r echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}
# user inputs

# specify the year you want a report for
report_year = 2023

# specify which species you want to include
filter_spp <- c("BCNH", "GBHE", "GREG", "SNEG", "DCCO")


# specify which sites, subregion or county you want to summarize
# you can specify multiple colonies by, eg:
# which_sites <- c("Bolinas", "Gold Hill")
# multiple counties, eg:
# which_sites <- c("ALA", "CCO")
# or a combination of colonies, counties and subregions, eg:
# which_sites <- c("MAR", "CSF", "Valley Ford")
# or just a single colony, subregion or county
which_sites <- c("Drakes Estero", "Hog Island", "Bear Valley", "Tree Farm")

# levels(hep$county)
# [1] "ALA" "CCO" "MAR" "NAP" "SAC" "SFR" "SOL" "SON"
# > levels(hep$subregion)
# [1] "CSF" "IEB" "NNC" "OUC" "PNM" "RUR" "SSF" "SUS"


```



```{r echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}


library(tidyverse)
library(RColorBrewer)
library(knitr)
library(flextable)
library(officer)
options(scipen = 999)
#source("C:/Users/scott.jennings/Documents/Projects/R_general/utility_functions/bird_utility_functions.R")
source("C:/Users/scott.jennings/OneDrive - Audubon Canyon Ranch/Projects/core_monitoring_research/HEP/HEP_data_work/HEP_code/HEP_utility_functions.R")


hepdata_location = "C:/Users/scott.jennings/OneDrive - Audubon Canyon Ranch/Projects/core_monitoring_research/HEP/HEP_data_work/HEP_data/HEPDATA.accdb"
# 2 data ----

# all these functions are in HEP_utility_functions.R
hep_start <- hep_from_access(hepdata_location)
hep_sites <- hep_sites_from_access(hepdata_location) 

hep_start <- hep_start %>% 
  append_as_hepdata()

hep <- hep_start %>% 
  clean_hep() %>% 
  filter(peakactvnsts >= 0) %>% 
  add_site_names() %>% 
  trim_hep_columns() %>% 
  cut_never_nested() %>% 
  mutate(site.name = as.character(site.name))
  
hep_sites <-  hep %>% 
  select(code, parent.code, site.name, parent.site.name, county, subregion) %>% 
  filter_all(any_vars(. %in% which_sites)) %>% 
  distinct()

hep_region_child <- hep %>%
  filter(code %in% hep_sites$code) %>% 
  filter(species %in% filter_spp) %>% 
  droplevels()

hep_region <- hep_region_child %>% 
  group_by(parent.site.name, parent.code, year, species) %>% 
  summarise(peakactvnsts = sum(peakactvnsts),
            focalnests = sum(focalnests),
            focfailure = sum(focfailure),
            numbervisits = sum(numbervisits),
            brd1 = sum(brd1),
            brd2 = sum(brd2),
            brd3 = sum(brd3),
            brd4 = sum(brd4),
            brd5 = sum(brd5),
            brd6 = sum(brd6)
            ) %>% 
  ungroup() %>% 
  rename(code = parent.code) %>% 
  mutate_if(is.numeric , replace_na, replace = 0)
 

#zspp_distinct = distinct(hep_region, species)

zspp = unname(unlist(distinct(hep_region, species)[,1]))
zsites <- unname(unlist(distinct(hep_region, parent.site.name)[,1]))

zsites_child <- unname(unlist(distinct(hep_region_child, site.name)[,1]))
```

```{r echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}
# new functions used just here


fill_all_spp_yrblocks <- function(znum_colonies) {
all_spp <- distinct(znum_colonies, species)
all_spp_yrblocks <- rbind(all_spp %>% mutate(year = "mean.all.previous.years"), 
                          all_spp %>% mutate(year = report_year)) 
znum_colonies <- znum_colonies %>% 
  full_join(., all_spp_yrblocks) %>% 
  replace(is.na(.), "0")
}       


# flextable formatting functions
flex_table_spp_header <- function(zflextable) {
  if("GREG" %in% zspp) {
    zflextable <- zflextable %>% set_header_labels(GREG_a = "Great Egret")
  } else {
    zflextable <- zflextable
  }
    if("GBHE" %in% zspp) {
    zflextable <- zflextable %>% set_header_labels(GBHE_a = "Great Blue Heron")
  } else {
    zflextable <- zflextable
  }
    if("BCNH" %in% zspp) {
    zflextable <- zflextable %>% set_header_labels(BCNH_a = "Black-crowned Night-heron")
  } else {
    zflextable <- zflextable
  }  
    if("SNEG" %in% zspp) {
    zflextable <- zflextable %>% set_header_labels(SNEG_a = "Snowy Egret")
  } else {
    zflextable <- zflextable
  }  
      if("CAEG" %in% zspp) {
    zflextable <- zflextable %>% set_header_labels(CAEG_a = "Cattle Egret")
  } else {
    zflextable <- zflextable
  }  
      if("DCCO" %in% zspp) {
    zflextable <- zflextable %>% set_header_labels(DCCO_a = "Double-crested Cormorant")
  } else {
    zflextable <- zflextable
  }  
  return(zflextable)
}

flex_table_merge_spp_cols <- function(zflextable) {
  if(any(grepl("DCCO", zflextable$body$col_keys)) == T) {
     zflextable_out <- zflextable %>%
  merge_at(i = 1, j = c("DCCO_a", "DCCO_b"), part = "header")
  } else {
    zflextable_out <- zflextable
  }
    if(any(grepl("GBHE", zflextable$body$col_keys)) == T) {
     zflextable_out <- zflextable_out %>%
  merge_at(i = 1, j = c("GBHE_a", "GBHE_b"), part = "header")
  } else {
    zflextable_out <- zflextable_out
  }
    if(any(grepl("GREG", zflextable$body$col_keys)) == T) {
     zflextable_out <- zflextable_out %>%
  merge_at(i = 1, j = c("GREG_a", "GREG_b"), part = "header")
  } else {
    zflextable_out <- zflextable_out
  }
    if(any(grepl("SNEG", zflextable$body$col_keys)) == T) {
     zflextable_out <- zflextable_out %>%
  merge_at(i = 1, j = c("SNEG_a", "SNEG_b"), part = "header")
  } else {
    zflextable_out <- zflextable_out
  }
    if(any(grepl("BCNH", zflextable$body$col_keys)) == T) {
     zflextable_out <- zflextable_out %>%
  merge_at(i = 1, j = c("BCNH_a", "BCNH_b"), part = "header")
  } else {
    zflextable_out <- zflextable_out
  }
    if(any(grepl("CAEG", zflextable$body$col_keys)) == T) {
     zflextable_out <- zflextable_out %>%
  merge_at(i = 1, j = c("CAEG_a", "CAEG_b"), part = "header")
  } else {
    zflextable_out <- zflextable_out
  }
return(zflextable_out)
}
  
 

flex_table_vert_borders <- function(zflextable) {
   zflextable_out <- zflextable
   if("DCCO_b" %in% zflextable$body$col_keys) {
     zflextable_out <- zflextable_out %>%
       border(j = "DCCO_b", border.right = fp_border(color = "black"), part = "body") %>% 
       border(j = "DCCO_b", i = 2, border.right = fp_border(color = "black"), part = "header")
  } else {
    zflextable_out <- zflextable_out
  }
     if("GBHE_b" %in% zflextable$body$col_keys) {
     zflextable_out <- zflextable_out %>%
       border(j = "GBHE_b", border.right = fp_border(color = "black"), part = "body") %>% 
       border(j = "GBHE_b", i = 2, border.right = fp_border(color = "black"), part = "header")
  } else {
    zflextable_out <- zflextable_out
  }
     if("GREG_b" %in% zflextable$body$col_keys) {
     zflextable_out <- zflextable_out %>%
       border(j = "GREG_b", border.right = fp_border(color = "black"), part = "body") %>% 
       border(j = "GREG_b", i = 2, border.right = fp_border(color = "black"), part = "header")
  } else {
    zflextable_out <- zflextable_out
  }
     if("SNEG_b" %in% zflextable$body$col_keys) {
     zflextable_out <- zflextable_out %>%
       border(j = "SNEG_b", border.right = fp_border(color = "black"), part = "body") %>% 
       border(j = "SNEG_b", i = 2, border.right = fp_border(color = "black"), part = "header")
  } else {
    zflextable_out <- zflextable_out
  }
     if("BCNH_b" %in% zflextable$body$col_keys) {
     zflextable_out <- zflextable_out %>%
       border(j = "BCNH_b", border.right = fp_border(color = "black"), part = "body") %>% 
       border(j = "BCNH_b", i = 2, border.right = fp_border(color = "black"), part = "header")
  } else {
    zflextable_out <- zflextable_out
  }
     if("CAEG_b" %in% zflextable$body$col_keys) {
     zflextable_out <- zflextable_out %>%
       border(j = "CAEG_b", border.right = fp_border(color = "black"), part = "body") %>% 
       border(j = "CAEG_b", i = 2, border.right = fp_border(color = "black"), part = "header")
  } else {
    zflextable_out <- zflextable_out
  }
return(zflextable_out)
}
```


```{r echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}
print("For anyone outputting this report, this is the list of subcolonies included in the report you just generated. It is included so you can confirm that the report really contains data for the colonies you want. You can delete this entire gray box before sharing the report.")
zsites_child
```

Summary of nest monitoring data for `r zsites` in `r report_year`.


Number of colonies monitored. Active colonies had at least 1 active nest during the report year. Innactive colonies had at least 1 active nests in at least 1 pervious year, but no active nests in the report year.
<br>

Note: Some colonies contain multiple species. This table indicates number of colonies where each species was nesting; the total number of colonies monitored was less than the sum of the Active row in this table (Active colonies table, below). Additionally, some colonies are comprised of multiple subcolonies; those are counted as single colonies here. 
```{r echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}
# number of colonies per year
colony_year <- hep_region %>% 
  mutate(active = ifelse(peakactvnsts > 0, 1, 0)) %>% 
  group_by(year, species, active) %>%
  summarise(num.colonies = n()) %>% 
  arrange(year, species, active) %>% 
  ungroup()
  
colony_year_wider <- colony_year %>% 
  spread(active, num.colonies) %>% 
  rename(inactive = 3, active = 4) %>% 
  dplyr::mutate(active = ifelse(is.na(active), 0, active),
         inactive = ifelse(is.na(inactive), 0, inactive),
         total = active + inactive) %>% 
  dplyr::select(year, species, total, active, inactive)

prev_years_num_colonies <- colony_year_wider %>% 
  filter(year < report_year) %>% 
  group_by(species) %>% 
  summarise_at(vars(total, active, inactive), funs(mean, sd)) %>% 
  mutate_if(is.numeric, round, 1) %>% 
  mutate(year = "mean.all.previous.years",
         total = paste(total_mean, " (", total_sd, ")", sep = ""),
         active = paste(active_mean, " (", active_sd, ")", sep = ""),
         inactive = paste(inactive_mean, " (", inactive_sd, ")", sep = "")) %>% 
  select(-contains("_mean"), -contains("_sd"))

cur_year_num_colonies <- colony_year_wider %>% 
  filter(year == report_year) 


num_colonies <- rbind(prev_years_num_colonies, cur_year_num_colonies) %>% 
  fill_all_spp_yrblocks()


num_colonies_longer <- num_colonies %>% 
  gather(colony.type, num.colonies, -species, -year) %>% 
  arrange(species) 

num_colonies_wide <- num_colonies_longer %>% 
  mutate(spp.year = ifelse(year == "mean.all.previous.years", paste(species, "b", sep = "_"), paste(species, "a", sep = "_"))) %>% 
  select(-species, -year) %>% 
  spread(spp.year, num.colonies) %>% 
  select(colony.type, matches(paste(zspp, collapse = "|"))) %>%
  #select(colony.type, contains("GREG"), contains("GBHE"), contains("BCNH"), contains("SNEG"), contains("CAEG")) %>%
  mutate(colony.type = sub("(.)", "\\U\\1", colony.type, perl=TRUE)) # capitalize first letter


#num_colonies_wide %>% 
#  data.frame() %>% 
#  flextable()

#kable(num_colonies_wide)
flextable(num_colonies_wide) %>% 
  flex_table_spp_header() %>% 
  set_header_labels(colony.type = "") %>% 
  flex_table_merge_spp_cols() %>% 
  add_header_row(values = c("Colony status", rep(c(report_year, "Mean of all previous years (SD)"), length(zspp))), top = FALSE) %>% 
  align_text_col(align = "left", header = TRUE, footer = TRUE) %>% 
  align(j = c(2:ncol(num_colonies_wide)), align = "center", part = "all") %>% 
  border_remove() %>% 
  # horizontal borders
  border(i = 1, border.top = fp_border(color = "black"), part = "header") %>% 
  border(i = 1, border.top = fp_border(color = "black"), part = "body") %>% 
  border(i = nrow(num_colonies_wide), border.bottom = fp_border(color = "black"), part = "body") %>% 
  # vertical borders
  flex_table_vert_borders() %>% 
  padding(padding = 2, part = "all") %>% 
  autofit() %>% 
  fit_to_width(max_width = 8)





```


<br>
Species nesting at each colony and number of monitoring visits.
```{r echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}
# number of colonies per year
colony_visits <- filter(hep_region, year == report_year) %>% 
  #filter(peakactvnsts > 0) %>% 
  group_by(parent.site.name) %>% 
  filter(numbervisits == max(numbervisits)) %>% 
  distinct(parent.site.name, numbervisits)

colony_species <- filter(hep_region, year == report_year) %>% 
  filter(peakactvnsts > 0) %>% 
  filter(species %in% filter_spp) %>% 
  mutate(common.name = translate_bird_names(species, "alpha.code", "common.name")) %>% 
  group_by(parent.site.name) %>% 
  summarise(species = paste(common.name, collapse = ", "))

colony_visits_species <- full_join(colony_visits, colony_species) %>% 
  rename(Colony = parent.site.name, Species = species) %>% 
  mutate(Species = ifelse(is.na(Species), paste("Not active", report_year), Species)) %>% 
  select(Colony, Species, numbervisits)


flextable(colony_visits_species) %>% 
  set_header_labels(numbervisits = "Number of visits") %>%
 align_text_col(align = "left", header = TRUE, footer = TRUE) %>% 
  align_nottext_col(align = "center", header = TRUE, footer = TRUE) %>% 
  border_remove() %>% 
  # horizontal borders
  border(i = 1, border.top = fp_border(color = "black"), part = "header") %>% 
  border(i = 1, border.top = fp_border(color = "black"), part = "body") %>% 
  border(i = nrow(colony_visits_species), border.bottom = fp_border(color = "black"), part = "body") %>% 
  padding(padding = 2, part = "all") %>% 
  autofit()
```

<br>
Total number of nests for each species, all colonies combined.
```{r echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}
# number of colonies per year
num_nests <- hep_region %>% 
  group_by(year, species) %>%
  summarise(total.nests = sum(peakactvnsts)) %>% 
  arrange(year, species) %>% 
  ungroup()
  

prev_years_num_nests <- num_nests %>% 
  filter(year < report_year) %>% 
  group_by(species) %>% 
  summarise_at(vars(total.nests), funs(mean, sd))%>% 
  mutate_if(is.numeric, round, 1) %>% 
  mutate(year = "mean.all.previous.years",
         total.nests = paste(mean, " (", sd, ")", sep = "")) %>% 
  select(-contains("mean"), -contains("sd"))

cur_year_num_nests <- num_nests %>% 
  filter(year == report_year)  %>% 
  mutate(year = report_year)

num_nests <- rbind(prev_years_num_nests, cur_year_num_nests) %>% 
  fill_all_spp_yrblocks()


num_nests_wide <- num_nests %>%
  mutate(spp.year = ifelse(year == "mean.all.previous.years", paste(species, "b", sep = "_"), paste(species, "a", sep = "_"))) %>% 
  select(-species, -year) %>% 
  spread(spp.year, total.nests) %>% 
  select(matches(paste(zspp, collapse = "|")))

flextable(num_nests_wide) %>% 
  flex_table_spp_header() %>% 
  flex_table_merge_spp_cols() %>% 
  add_header_row(values = c(rep(c(report_year, "Mean of all\nprevious years (SD)"), length(zspp))), top = FALSE) %>% 
  align(align = "center", part = "all") %>% 
  border_remove() %>% 
  # horizontal borders
  border(i = 1, border.top = fp_border(color = "black"), part = "header") %>% 
  border(i = 1, border.top = fp_border(color = "black"), part = "body") %>% 
  border(i = nrow(num_nests_wide), border.bottom = fp_border(color = "black"), part = "body") %>% 
  # vertical borders
  flex_table_vert_borders() %>% 
  padding(padding = 2, part = "all") %>% 
  autofit() %>% 
  fit_to_width(max_width = 8)


```




<br>
Average brood size for each species, all colonies combined.
```{r echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}


 

brsz_prv_yrs <- hep_region %>% 
  filter(year < report_year) %>% 
  calc_mean_brood_size_lump_yr_col() %>% 
  filter(!(is.na(mean.brd.size))) %>% 
  distinct(species, mean.brd.size, sd.brd.size, brd.n.nests) %>% 
  mutate(year = "mean.all.previous.years")

brsz_report_yr <- hep_region %>% 
  filter(year == report_year) %>% 
  calc_mean_brood_size_lump_yr_col() %>% 
  filter(!(is.na(mean.brd.size))) %>% 
  distinct(species, mean.brd.size, sd.brd.size, brd.n.nests) %>% 
  mutate(year = report_year)

   


brood_size <- rbind(brsz_report_yr, brsz_prv_yrs) %>% 
  fill_all_spp_yrblocks()

brood_size_all <- expand.grid(species = brood_size$species, year = brood_size$year) %>% 
  distinct() %>% 
  full_join(., brood_size) %>% 
  mutate(across(contains("brood.size"), ~str_replace("NaN", "-", brood.size)),
         across(contains("brood.size"), ~ifelse(is.na(.), "-", .)))


brood_size_wide <- brood_size %>% 
  mutate(spp.year = ifelse(year == "mean.all.previous.years", paste(species, "b", sep = "_"), paste(species, "a", sep = "_")),
         tab.val = paste(round(mean.brd.size, 2), " (", round(sd.brd.size, 2), ")", sep = "")) %>%  
  select(spp.year, tab.val) %>% 
  spread(spp.year, tab.val) %>% 
  select(matches(paste(zspp, collapse = "|")))

flextable(brood_size_wide) %>% 
  flex_table_spp_header() %>% 
  flex_table_merge_spp_cols() %>% 
  add_header_row(values = c(rep(c(paste(report_year, "(SD)"), "Mean of all\nprevious years (SD)"), length(.$body$col_keys)/2)), top = FALSE) %>% 
  align(align = "center", part = "all") %>% 
  border_remove() %>% 
  # horizontal borders
  border(i = 1, border.top = fp_border(color = "black"), part = "header") %>% 
  border(i = 1, border.top = fp_border(color = "black"), part = "body") %>% 
  border(i = nrow(brood_size_wide), border.bottom = fp_border(color = "black"), part = "body") %>% 
  # vertical borders
  flex_table_vert_borders() %>%  
  padding(padding = 2, part = "all") %>% 
  autofit() %>% 
  fit_to_width(max_width = 8)

```

